A modelling strategy to improve cacao quality and productivity

### 2021-07-30 papers/model/paper_model.tex
introduction cocoa models in progress

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{As.}

@@

Cocoa ( \textit{Theobroma cacao }L. ) is a perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}. Only the 5\% of the world cocoa yield is desalinated for Fine-cocoa production due to the low productivity of  the traditional crop management \citep{argout2011}. Cacao plant member of the Malvaceae (formerly Sterculiaceae)  botanical family, is grown for its fruits, known as cacao pods.  \citep{ Niemenak2010, suarez2021}.

Cocoa has  a  long  tradition  in  Colombia  where  cocoa  is  widely  consumed  as  a  beverage. It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019} (Abbott et al. 2018).

In this work we used the SIMPLE crop model \citep{Zao2019simple} for three reasons: 1: That it is very comprehensively described in the original paper. 2: That the code was available in R for initial trials and 3: That it had already been successfully fitted to perennial crops in south America. Overall, the model simulates crop development, growth and yield, and predict the maturation day when the friut is ready to harvest. It includes 13 parameters to specify a crop type, with four of these for cultivar characteristics. Commonly available inputs that are required for the model include daily weather data, crop management, and soil water holding parameters and key dates. Malvaceae such as cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton} wich is modeled in SIMPLE model \citep{Zao2019simple}. 
 The aim of this study is predict the best time for harvest cocoa fruits.

@@

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{ The aim of this study is predict the best time for harvest cocoa fruits.}

@@

Crop models represent a quantitative assumption of plant growth depending on sunlight interception efficiency values and climate data supported by a large amount of empirical and ground data. \citep{Reynolds2018}. Physiological crop models have shown to be very useful tool for provided agronomical advices and improvements of the cropping systems of annual crops mainly. Recently crop modelling studies are focusing on  perennial crops  production \citep{zuidema2005, Zao2019simple, Bai2020, Romero2021}. However, the information reported  is be less than for annual crops due to the lack of field data available , relatively high research costs and the difficulties of accumulated errors in long-term simulations \citep{zuidema2005}. For cacao there  approaches  to predict yield mainly  using algorithms of machine learning \citep{lamos2020} and just one simulate cocoa performance as physiological production model "SUCROS-cocoa" \citep{zuidema2005}.

Regarding cocoa crops, the yield prediction has been approached from different perspectives. Corrales et al. [11] predicted the cocoa yield level in Santander. The authors evaluated the daily average temperature, daily relative humidity, and total
daily precipitations rate, using ten different algorithms implemented in WEKA software. For them, Random Forest was the algorithm that generates the best model in order to classify cocoa yield levels. Other studies [12-13] evaluate the yield using linear regression models, ANOVA and mechanistic models like SUCROS, finding that climatic conditions (such as temperature, radiation and rainfall) are the most
critical in the cocoa productivity \citep{lamos2020}.

Cocoa ( \textit{Theobroma cacao }L. ) is an important worldwide perennial tropical crop endemic to the South American rainforests \citep{zuidema2005, motamayor2002, argout2011, Rodriguez2019}. Cacao plant member of the Malvaceae (formerly Sterculiaceae)  botanical family, is grown for its fruits, known as cacao pods.  \citep{ Niemenak2010, suarez2021}. Only the 5\% of the world cocoa yield is desalinated for Fine-cocoa production due to the low productivity of  the traditional crop management \citep{argout2011}.  In Colombia, cocoa  is  traditionally  consumed  as  a  beverage. It is one of the crops promoted by the Colombian government in the social and agricultural development  programs aimed at favouring peace in post-conflict regions \citep{Rodriguez2019, Abbott2019} as cocoa is grown by approximately 52.000 \citep{Gutierrez2020} and 98\% of production being carried out by small and medium-sized producers \citep{Garcia2014, Escobar2020}. Colombia registered an increase of 3.750 tons in production in 2020 compared to the previous year \citep{lamos2020}.

Although Colombian cocoa has the potential to be in the high value markets for fine flavour \citep{Escobar2020}. However, it is still not widely produced as the lack of adoptions of technologies by the traditional farmers. They empirically harvest after 5 or 6 months after flowering date, hence they ferment cocoa beans without considering the  quality of the seeds at the harvest time. This produce heterogeneous characteristics between each fermentation batch diminishing the quality of cocoa final product \citep{Escobar2021}. To identify the best moment to harvest is important to consider physiological responses affected by climate variables such as rain, solar radiation and wind.   Crop simulation models (CSM) are mathematical calculations include in software tools that simulated the behaviour of crop systems \citep{Oteng-Darko2013}. For cocoa in Colombia, physiological simulation models may be valuable to identify the best moment to harvest cocoa considering variable weather conditions, soil types and cropping systems.

In this paper, we present a physiological parametrization of SIMPLE crop model for cocoa to predict the best harvest time  and yield production. We used the SIMPLE crop model \citep{Zao2019simple} for three reasons: 1: That it is very comprehensively described in the original paper. 2: That the code was available in R for initial trials and 3: That it had already been successfully fitted to perennial crops in south America. Overall, the model simulates crop development, growth and yield, and predict the maturation day when the fruit is ready to harvest. It includes 13 parameters to specify a crop type, with four of these for cultivar characteristics. Commonly available inputs that are required for the model include daily weather data, crop management, and soil water holding parameters and key dates. Malvaceae such as cotton \textit{ Gossypium hirstium} \citep{Nix2017cotton} wich is modeled in SIMPLE model \citep{Zao2019simple}.

@@

Evaluating the level of knowledge of producers regarding cocoa crop management, the harvest was in the group of activities that presented the lowest level of knowledge on the part of the producers according to the general averages \citep{Gutierrez2020}.
